Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-03 paper_jasa.tex
small editorial changes

\begin{dfn}\label{dfn:test}
The \textbf{visual test}, $V_{\theta}$  of size $\alpha$, is defined to be 
\begin{itemize}\itemsep-3pt
\item \textbf{Reject} $H_0$ if out of $K$ observers $x_{\alpha}$ of them correctly identify the actual data plot in the lineup, and
\item \textbf{Fail to reject} $H_0$  otherwise. %if the observer selects a null plot.
\end{itemize}
where $x_{\alpha}$ is such that $P(X \ge x_{\alpha}) < \alpha$. 
\end{dfn}

Associated with any test is the risk of making a Type I error, or Type II error, which for visual inference are defined as follows:

\begin{dfn}\label{dfn:error}
The \textbf{Type I error} is the probability of rejecting $H_0$ when it is true, the probability for that is $P(X \ge x_{\alpha})$, which is controlled by $\alpha$.
% identifying the actual data plot, when $H_0$ is true, which is $1/m$, and is thus controlled by the number of null plots. 
\textbf{Type II error} is the probability of failing to identify the actual data plot, when $H_0$ is not true, $P( X <  x_{\alpha})$.
\end{dfn}

\green{Oh, we are using observed and observer, which might be confusing. May need to change observed data plot to something else. Will think about this. Will use ``actual'' for now.}

\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$, with an exact $x_{\alpha}$. For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$. It can be set to be smaller, even arbitrarily small, by increasing $K$. Type II error is harder to calculate, as is usually the case. In visual inference, individual abilities need to be incorporated to calculate Type II error. This is not necessary for Type I error, because it is simply the chance that the observer picks the actual data plot, by chance. For Type II error, we need to calculate the chance that the observer sees that actual data plot as different, when it really is different. This involves understanding the individual's visual skills. Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise, where $p_i$ is the probability that individual $i$ picks out the actual data plot from the lineup. If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}_{K, p}$. We can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.

\hh{If there is evidence for individual skills influencing the probability $p_i$, then $X_i \sim Binom_{1, p+i}$ and $X$ is a sum of independent Bernoulli random variables with different success rates $p_i$. This makes the distribution of $X$ a Poisson-Binomial by definition, see \citet{butler93} for details. estimating $p_i$ will be one of the main concerns for the remainder of the paper.}

>>>>

\begin{dfn}\label{dfn:test}
The \textbf{visual test}, $V_{\theta}$  of size $\alpha$, is defined to be 
\begin{itemize}\itemsep-3pt
\item \textbf{Reject} $H_0$ if out of $K$ observers at least $x_{\alpha}$ correctly identify the actual data plot in the lineup, and
\item \textbf{Fail to reject} $H_0$  otherwise. 
\end{itemize}
where $x_{\alpha}$ is such that $P(X \ge x_{\alpha}) < \alpha$. 
\end{dfn}

\noindent
Associated with any test is the risk of  Type I or II errors, which for visual inference are defined as follows:

\begin{dfn}\label{dfn:error}
The \textbf{Type I error} associated with visual test $V_{\theta}$ is the probability of rejecting $H_0$ when it is true; the probability for that is $P(X \ge x_{\alpha})$, which is controlled by $\alpha$.
\textbf{Type II error} is the probability of failing to identify the actual data plot, when $H_0$ is not true, $P( X <  x_{\alpha})$.
\end{dfn}

%\green{Oh, we are using observed and observer, which might be confusing. May need to change observed data plot to something else. Will think about this. Will use ``actual'' for now.} \hh{I think that change is done}

\noindent Because $X$ takes only discrete values we can not always control exactly for $\alpha$, with an exact $x_{\alpha}$. For example, when there is only one observer, $1/m$ is the minimal value at which we can set $\alpha$. It can be set to be smaller, even arbitrarily small, by increasing $K$, the number of observers. 
Type II error is harder to calculate, as is usually the case. In visual inference, individual abilities need to be incorporated to calculate Type II error. This is not necessary for Type I error, because it is simply the chance that the observer picks the actual data plot, by chance. For Type II error, we need to calculate the chance that the observer sees that actual data plot as different, when it really is different. This involves understanding the individual's visual skills. Thus, let $X_i$ be a binary random variable with $X_i = 1$, if individual $i (=1, \dots , K)$ identifies the actual data plot from the lineup, and $X_i = 0$ otherwise, where $p_i$ is the probability that individual $i$ picks out the actual data plot from the lineup. If all individuals have the same ability, with the probability, $p$, for picking out the actual data plot, then $X = \sum_i X_i$ has distribution $\text{Binom}_{K, p}$, and we can estimate $p$ by $\hat{p} = x/K$, where $x$ is the number of observers (out of $K$), who pick out the actual data plot from the lineup.

\hh{If there is evidence for individual skills influencing the probability $p_i$, then $X_i \sim Binom_{1, p+i}$ and $X$ is a sum of independent Bernoulli random variables with different success rates $p_i$. This makes the distribution of $X$ a Poisson-Binomial by definition, see \citet{butler93} for details. Estimating $p_i$ will be one of the main concerns for the remainder of the paper.}
